Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-01-30 draft.tex
explicit potential in main theorem

>>>>

\section{....}

\begin{proof}
    We introduce the reflection mapping $\Xi_{1} : T_{m} \rightarrow \Makoto_{m-1}$ of Proposition~\ref{proposition:starreflection}. 
    We define 
    \begin{align*}
        w'_{m} :=  \Xi_{1}^{\ast} w_{m-1} \in W^{p}\Alt^{k}(T_{m}).
    \end{align*}
    By construction, $w'_{m} \in W^{p}\Alt^{k}(T_{m})$ with 
    \begin{align*}
        \trace_F w_{m-1} = \trace_F w'_{m}
        . 
    \end{align*}
    There exists $w''_{m} \in W^{p}\Alt^{k}(T_{m})$ satisfying 
    \begin{align*}
        \cartan w''_{m} = \cartan u - \cartan w'_{m}, 
        \quad 
        \trace_{\Gamma_{m}} w''_{m} = 0.
    \end{align*}
    Indeed, such $w''_{m}$ exists because 
    \begin{align*}
        \trace_{\Gamma_{m}}( \cartan u - \cartan w'_{m} ) 
        &= 
        \trace_{\Gamma_{m}} \cartan u - \trace_{\Gamma_{m}} \cartan w'_{m}
        \\&= 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w'_{m}
        = 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w_{m-1}
        = 
        0
        .
    \end{align*}
    Setting $w_{m} := w_{m-1}$ over $X_{m-1}$ and $w_{m} := w'_{m} + w''_{m}$ over $T_{m}$, 
    we find 
    \begin{gather*}
        \cartan w_{m} = \cartan w'_{m} + \cartan w''_{m} = \cartan w'_{m} + \cartan u - \cartan w'_{m} = \cartan u,
        \\
        \trace_{\Gamma_{m}} w_{m} = \trace_{\Gamma_{m}} w'_{m} = \trace_{\Gamma_{m}} w_{m-1}.
    \end{gather*}
    That means that $w_{m} \in W^{p}\Alt^{k}(X_{m})$ with $\cartan w_{m} = \cartan u$ over $X_{m}$.

    We want to estimate norms. By the triangle inequality, 
    \begin{align*}
        \| w_{m} \|_{L^{p}(T_{m})}
        \leq  
        \| w'_{m} \|_{L^{p}(T_{m})}
        + 
        \| w''_{m} \|_{L^{p}(T_{m})}
        .
    \end{align*}
    Due to Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative},
    we can assume that $w''_{m}$ satisfies 
    \begin{align*}
        \| w''_{m} \|_{L^{p}(T_{m})} 
        &
        \leq 
        C_{\PF,T_{m},\Gamma_{m},k,p} \| \cartan u - \cartan w'_{m} \|_{L^{p}(T_{m})}
        \\&
        \leq 
        C_{\PF,T_{m},\Gamma_{m},k,p} \| \cartan u      \|_{L^{p}(T_{m})} 
        + 
        C_{\PF,T_{m},\Gamma_{m},k,p} \| \cartan w'_{m} \|_{L^{p}(T_{m})} 
        .
    \end{align*}
    Proposition~\ref{proposition:pullbackestimate} and Proposition~\ref{proposition:starreflection} now show 
    % \begin{align*}
        % \| w'_{m} \|_{L^{p}(T_{m})}
        % &\leq 
        % \| \Jacobian \Xi_{1} \|_{L^{\infty}(T_{m})}^{k}
        % \|\det \Jacobian \Xi_{1}^{-1}  \|_{L^{\infty}(\Makoto_{m-1})}^{\frac 1 p} 
        % \cdot 
        % \| w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        % ,
        % \\
        % \| \cartan w'_{m} \|_{L^{p}(T_{m})}
        % &\leq 
        % \| \Jacobian \Xi_{1} \|_{L^{\infty}(T_{m})}^{k+1}
        % \|\det \Jacobian \Xi_{1}^{-1}  \|^{\frac 1 p}_{L^{\infty}(\Makoto_{m-1})} 
        % \cdot 
        % \| \cartan w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        % .
    % \end{align*}
    \begin{align*}
        \| w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        \Cfivetrafo{n}{k}{\ell}{p}(\calT)
        % \Cfive{n}{k}(\calT)
        % \Cfiveprime{n}{k}(\calT)^{k-1}
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p}
        \cdot 
        \| w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        ,
        \\
        \| \cartan w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        \Cfivetrafo{n}{k+1}{\ell}{p}(\calT)
        % \Cfive{n}{k}{\ell}(\calT)
        % \Cfiveprime{n}{k}(\calT)^{k}
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p}
        \cdot 
        \| \cartan w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        .
    \end{align*}
    We have assumed that $\cartan w_{m-1} = \cartan u_{|X_{m-1}}$. 
    The existence of $w \in W^{p}\Alt^{k}(\Omega)$ satisfying the recursive estimate follows. 
\end{proof}

### 2025-01-30 main.tex
explicit potential in main theorem

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    % Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable $n$-dimensional triangulation $\calT$.
    Let $\calT$ be a shellable $n$-dimensional triangulation, and let the domain $\Omega \subseteq \bbR^{n}$ be the interior of the underlying set of $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling of $\calT$.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$, where $1 \leq p \leq \infty$, 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$
    and such that the following estimates hold:
    \begin{gather*}
        \| w \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}
        ,
    \end{gather*}
    and for each $T_{m} \in \calT$, $1 \leq m \leq M$ we have the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfivetrafo{n}{k+1}{\ell}{p}(\calT)
            % \Cfive{n}{k}(\calT)
            % \Cfiveprime{n}{k}(\calT)^{k} 
            % \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(\Makoto_{m-1})}
        \right)
        %\\&\qquad
        + 
        \Cfivetrafo{n}{k}{\ell}{p}(\calT)
        % \Cfive{n}{k}(\calT) 
        % \Cfiveprime{n}{k}(\calT)^{k-1} 
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(\Makoto_{m-1})}
        ,
    \end{align*}
    where $0 \leq \ell < n$ is such that $T_{m}$ has $n - \ell$ faces in common with the previous simplices. 
\end{theorem}
\begin{proof}
    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(X_{m-1})$ 
    with $\cartan w_{m-1} = \cartan u$ over $X_{m-1}$. 
    By assumption, $T_{m}$ and $X_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex $S_{m}$, 
    and we let $\Makoto_{m-1} \subseteq X_{m-1}$ be the complement of $T_{m}$ in that newly completed star. 
    Write $\ell$ for the dimension of $S_{m}$.

    We introduce the reflection mapping $\Xi_{1} : T_{m} \rightarrow \Makoto_{m-1}$ of Proposition~\ref{proposition:starreflection}. 
    We define 
    \begin{align*}
        w'_{m} :=  \Xi_{1}^{\ast} w_{m-1} \in W^{p}\Alt^{k}(T_{m}).
    \end{align*}
    By construction, $w'_{m} \in W^{p}\Alt^{k}(T_{m})$ with 
    \begin{align*}
        \trace_F w_{m-1} = \trace_F w'_{m}
        . 
    \end{align*}
    There exists $w''_{m} \in W^{p}\Alt^{k}(T_{m})$ satisfying 
    \begin{align*}
        \cartan w''_{m} = \cartan u - \cartan w'_{m}, 
        \quad 
        \trace_{\Gamma_{m}} w''_{m} = 0.
    \end{align*}
    Indeed, such $w''_{m}$ exists because 
    \begin{align*}
        \trace_{\Gamma_{m}}( \cartan u - \cartan w'_{m} ) 
        &= 
        \trace_{\Gamma_{m}} \cartan u - \trace_{\Gamma_{m}} \cartan w'_{m}
        \\&= 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w'_{m}
        = 
        \cartan \trace_{\Gamma_{m}} u - \cartan \trace_{\Gamma_{m}} w_{m-1}
        = 
        0
        .
    \end{align*}
    Setting $w_{m} := w_{m-1}$ over $X_{m-1}$ and $w_{m} := w'_{m} + w''_{m}$ over $T_{m}$, 
    we find 
    \begin{gather*}
        \cartan w_{m} = \cartan w'_{m} + \cartan w''_{m} = \cartan w'_{m} + \cartan u - \cartan w'_{m} = \cartan u,
        \\
        \trace_{\Gamma_{m}} w_{m} = \trace_{\Gamma_{m}} w'_{m} = \trace_{\Gamma_{m}} w_{m-1}.
    \end{gather*}
    That means that $w_{m} \in W^{p}\Alt^{k}(X_{m})$ with $\cartan w_{m} = \cartan u$ over $X_{m}$.

    We want to estimate norms. By the triangle inequality, 
    \begin{align*}
        \| w_{m} \|_{L^{p}(T_{m})}
        \leq  
        \| w'_{m} \|_{L^{p}(T_{m})}
        + 
        \| w''_{m} \|_{L^{p}(T_{m})}
        .
    \end{align*}
    Due to Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative},
    we can assume that $w''_{m}$ satisfies 
    \begin{align*}
        \| w''_{m} \|_{L^{p}(T_{m})} 
        &
        \leq 
        C_{\PF,T_{m},\Gamma_{m},k,p} \| \cartan u - \cartan w'_{m} \|_{L^{p}(T_{m})}
        \\&
        \leq 
        C_{\PF,T_{m},\Gamma_{m},k,p} \| \cartan u      \|_{L^{p}(T_{m})} 
        + 
        C_{\PF,T_{m},\Gamma_{m},k,p} \| \cartan w'_{m} \|_{L^{p}(T_{m})} 
        .
    \end{align*}
    Proposition~\ref{proposition:pullbackestimate} and Proposition~\ref{proposition:starreflection} now show 
    % \begin{align*}
        % \| w'_{m} \|_{L^{p}(T_{m})}
        % &\leq 
        % \| \Jacobian \Xi_{1} \|_{L^{\infty}(T_{m})}^{k}
        % \|\det \Jacobian \Xi_{1}^{-1}  \|_{L^{\infty}(\Makoto_{m-1})}^{\frac 1 p} 
        % \cdot 
        % \| w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        % ,
        % \\
        % \| \cartan w'_{m} \|_{L^{p}(T_{m})}
        % &\leq 
        % \| \Jacobian \Xi_{1} \|_{L^{\infty}(T_{m})}^{k+1}
        % \|\det \Jacobian \Xi_{1}^{-1}  \|^{\frac 1 p}_{L^{\infty}(\Makoto_{m-1})} 
        % \cdot 
        % \| \cartan w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        % .
    % \end{align*}
    \begin{align*}
        \| w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        \Cfivetrafo{n}{k}{\ell}{p}(\calT)
        % \Cfive{n}{k}(\calT)
        % \Cfiveprime{n}{k}(\calT)^{k-1}
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p}
        \cdot 
        \| w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        ,
        \\
        \| \cartan w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        \Cfivetrafo{n}{k+1}{\ell}{p}(\calT)
        % \Cfive{n}{k}{\ell}(\calT)
        % \Cfiveprime{n}{k}(\calT)^{k}
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p}
        \cdot 
        \| \cartan w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        .
    \end{align*}
    We have assumed that $\cartan w_{m-1} = \cartan u_{|X_{m-1}}$. 
    The existence of $w \in W^{p}\Alt^{k}(\Omega)$ satisfying the recursive estimate follows. 
\end{proof}

>>>>

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    % Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable $n$-dimensional triangulation $\calT$.
    Let $\calT$ be a shellable $n$-dimensional triangulation, and let the domain $\Omega \subseteq \bbR^{n}$ be the interior of the underlying set of $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling of $\calT$.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$, where $1 \leq p \leq \infty$, 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$
    and such that the following estimates hold:
    \begin{gather*}
        \| w \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}
        ,
    \end{gather*}
    and for each $T_{m} \in \calT$, $1 \leq m \leq M$ we have the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfivetrafo{n}{k+1}{\ell}{p}(\calT)
            % \Cfive{n}{k}(\calT)
            % \Cfiveprime{n}{k}(\calT)^{k} 
            % \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(\Makoto_{m-1})}
        \right)
        %\\&\qquad
        + 
        \Cfivetrafo{n}{k}{\ell}{p}(\calT)
        % \Cfive{n}{k}(\calT) 
        % \Cfiveprime{n}{k}(\calT)^{k-1} 
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(\Makoto_{m-1})}
        ,
    \end{align*}
    where $0 \leq \ell < n$ is such that $T_{m}$ has $n - \ell$ faces in common with the previous simplices. 
\end{theorem}
\begin{proof}
    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(X_{m-1})$ with $\cartan w_{m-1} = \cartan u$ over $X_{m-1}$. 
    This is already true for $m = 1$.
    By assumption, $T_{m}$ and $X_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex $S_{m}$, 
    and we let $\Makoto_{m-1} \subseteq X_{m-1}$ be the complement of $T_{m}$ in that newly completed star. 
    Write $\ell$ for the dimension of $S_{m}$.

    We now define, as in~\cite[Equations~{\char`\(}5.12{\char`\)} and~{\char`\(}5.14{\char`\)}]{ern2020stable} or~\cite[equations~{\char`\(}6.7{\char`\)} and~{\char`\(}6.9{\char`\)}]{Chaum_Voh_p_rob_3D_H_curl_24}, the field $\widetilde w_{m} \in W^{p}\Alt^{k}(T_{m})$ via 
    \begin{align} \label{equation:w''_ext_der}
        {\widetilde w}_{m} := u_{|T_m} + \Xi_{1}^{\ast}\left( (w_{m-1} - u)_{|\Makoto_{m-1} } \right).
    \end{align}
    One verifies that 
    \begin{align*}
        \trace_F {\widetilde w}_{m} 
        &
        = 
        \trace_F u_{|T_m} + \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \trace_F \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \trace_F u_{|T_m} + \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \trace_F \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} }
        \\&
        = 
        \trace_F w_{m-1|\Makoto_{m-1}}
        .
    \end{align*}
    Moreover, 
    \begin{align*}
        \cartan {\widetilde w}_{m} 
        &
        = 
        \cartan u_{|T_m} + \cartan \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \cartan \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \cartan u_{|T_m} + \Xi_{1}^{\ast} \cartan w_{m-1|\Makoto_{m-1} } - \Xi_{1}^{\ast} \cartan u_{|\Makoto_{m-1} } 
        = 
        \cartan u_{|T_m}
        .
    \end{align*}
    Setting $w_{m} := w_{m-1}$ over $X_{m-1}$ and $w_{m} := {\widetilde w}_{m} + w''_{m}$ over $T_{m}$, 
    we thus verify $w_{m} \in W^{p}\Alt^{k}(X_{m})$ with $\cartan w_{m} = \cartan u$ over $X_{m}$.

    We want to estimate norms. 
    By construction,
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        \leq  
        \| \Xi_{1}^{\ast} w_{|\Makoto_{m-1}} \|_{L^{p}(T_{m})}
        + 
        \| u - \Xi_{1}^{\ast} u_{|\Makoto_{m-1}} \|_{L^{p}(T_{m})}
        .
    \end{align*}
    Due to Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative}, 
    which applies since $\trace_{\Gamma_{m}} \left( u_{|T_m} - \Xi_{1}^{\ast} u_{|\Makoto_{m-1}} \right) = 0$, 
    we have
    \begin{align*}
        \| u - \Xi_{1}^{\ast} u_{|\Makoto_{m-1}} \|_{L^{p}(T_{m})} 
        &
        \leq 
        C_{\PF,T_{m},\Gamma_{m-1},k,p} 
        \left( 
            \| \cartan u \|_{L^{p}(T_{m})} + \| \cartan \Xi_{1}^{\ast} u_{|\Makoto_{m-1}} \|_{L^{p}(T_{m})} 
        \right)
        \\&
        \leq 
        C_{\PF,T_{m},\Gamma_{m-1},k,p} 
        \left( 
            \| \cartan u \|_{L^{p}(T_{m})} + \| \Xi_{1}^{\ast} \cartan u_{|\Makoto_{m-1}} \|_{L^{p}(T_{m})} 
        \right)
        .
    \end{align*}
    Proposition~\ref{proposition:pullbackestimate} and Proposition~\ref{proposition:starreflection} now show 
    \begin{align*}
        \| \Xi_{1}^{\ast} w_{|\Makoto_{m-1}} \|_{L^{p}(T_{m})}
        &\leq 
        \Cfivetrafo{n}{k}{\ell}{p}(\calT)
        % \Cfive{n}{k}(\calT)
        % \Cfiveprime{n}{k}(\calT)^{k-1}
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p}
        \| w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        ,
        \\
        \| \Xi_{1}^{\ast} \cartan u_{|\Makoto_{m-1}} \|_{L^{p}(T_{m})}
        &\leq 
        \Cfivetrafo{n}{k+1}{\ell}{p}(\calT)
        % \Cfive{n}{k}{\ell}(\calT)
        % \Cfiveprime{n}{k}(\calT)^{k}
        % \Cdetsechs{n}{k}(\calT)^{\frac 1 p}
        \| \cartan w_{m-1} \|_{L^{p}(\Makoto_{m-1})}
        .
    \end{align*}
    We have assumed that $\cartan w_{m-1} = \cartan u_{|X_{m-1}}$. 
    The existence of $w \in W^{p}\Alt^{k}(\Omega)$ satisfying the recursive estimate follows. 
\end{proof}
